---
title: "Fit the Bayesian RMR regression"
output: html_notebook
---

This file fits and saves the Bayesian RMR regression model. It is intended to be run before any of the main figuree generation files which load and depend on objects created and saved locally by this file. 

This file is only intended to be run once in order to generate the fitted model objects and defined parameters to be used in the subsequent figure generation files. 

## Setup

```{r}

# --- libraries
library(phytools)        # pylogenetic analysis
library(nlme)            # pgls analysis
library(plyr)            # organise dataset
library(yarrr)           # transparent color
library(viridis)         # color palettes
library(brms)            # fit baysian models
library(parallel)        # run models in parallel
library(rstan)           # run bayes
library(bayestestR)      # Bayesian p-values
library(geiger)          # covariance matrix
library(predictFishRMR)
library(MatchIt)
```

## Import data

```{r}
# --- open data
# rmr_data <- read.csv("MandT_indiv_whalefixed_raw.csv")
#rmr_data <- read.csv("MandT_indiv_whalefixed_raw_CORE.csv")
rmr_data <- read.csv("RMR_data_final.csv")

tree <- read.tree("fish_shark_combined.tre")
```

## Process Tree data

```{r}

# Compare species between dataset and phylogeny
unique(rmr_data$spp)
tree$tip.label

tree$tip.label <- gsub("_", " ", tree$tip.label)

tree$tip.label[tree$tip.label == "Pseudocaranx dentex -formaly Longirostrum delicatissimus"] <- "Pseudocaranx dentex (formaly Longirostrum delicatissimus) "
tree$tip.label[tree$tip.label == "Ranzania laevis- Ostracion boops stage"] <- "Ranzania laevis, Ostracion boops stage"
tree$tip.label[tree$tip.label == "Coregonus lavaretus -larvae"] <- "Coregonus lavaretus (larvae)"
tree$tip.label[tree$tip.label == "Sardinops sagax -formaly S. caerulea"] <- "Sardinops sagax (formaly S. caerulea)"
tree$tip.label[tree$tip.label == "Ophichthus gomesii- leptocephalus larvae"] <- "Ophichthus gomesii, leptocephalus larvae"
tree$tip.label[tree$tip.label == "Paraconger caudilimbatus- leptocephalus larvae"] <- "Paraconger caudilimbatus, leptocephalus larvae"
tree$tip.label[tree$tip.label == "Ariosoma balearicum- leptocephalus larvae"] <- "Ariosoma balearicum, leptocephalus larvae"
tree$tip.label[tree$tip.label == "Gymnothorax saxicola- leptocephalus larvae"] <- "Gymnothorax saxicola, leptocephalus larvae"

rmr_data$spp[rmr_data$spp == "Negaprion brevirostris\xa0"] <- "Negaprion brevirostris"
rmr_data$spp[rmr_data$spp == "Mustelus antarcticus\xa0"] <- "Mustelus antarcticus"
rmr_data$spp[rmr_data$spp == "A. fimbria"] <- "Anoplopoma fimbria"
rmr_data$spp[rmr_data$spp == "Oncorhynchus\xa0tshawytscha"] <- "Oncorhynchus tshawytscha"
rmr_data$spp[rmr_data$spp == "Trachurus capensis\xa0"] <- "Trachurus capensis"
rmr_data$spp[rmr_data$spp == "Hypoatherina sp"] <- "Hypoatherina sp."
rmr_data$spp[rmr_data$spp == "Monacanthidae sp"] <- "Monacanthidae"
rmr_data$spp[rmr_data$spp == "Monacanthidae sp."] <- "Monacanthidae"
rmr_data$spp[rmr_data$spp == "Ambassis sp"] <- "Ambassis sp."
rmr_data$spp[rmr_data$spp == "Caulophrynidae sp."] <- "Caulophrynidae"
rmr_data$spp[rmr_data$spp == "Scomberesocidae sp."] <- "Scomberesocidae"
rmr_data$spp[rmr_data$spp == "Carcharodon carcharius"] <- "Carcharodon carcharias"
rmr_data$spp[rmr_data$spp == "Thunnys maccoyyi"] <- "Thunnus maccoyii"

# Especies en el dataset que faltan en el árbol
missing_in_tree <- setdiff(unique(rmr_data$spp), tree$tip.label)
missing_in_tree

# Especies en el árbol que no están en el dataset
extra_in_tree <- setdiff(tree$tip.label, unique(rmr_data$spp))
extra_in_tree

# Podar el árbol para eliminar especies extra
tree <- drop.tip(tree, extra_in_tree)

# Especies comunes entre ambos (dataset y árbol)
common_species <- intersect(unique(rmr_data$spp), tree$tip.label)
common_species

# Check if all dataset species are present in phylogeny
all_match <- all(sort(unique(rmr_data$spp)) %in% sort(tree$tip.label))
all_match  # TRUE if all dataset species are included in the tree

# --- housekeeping
rmr_data$method <- factor(rmr_data$method)
rmr_data$therm <- factor(rmr_data$therm)
rmr_data <- rmr_data[complete.cases(rmr_data),]
```

Summarise the data

```{r}

# add log(mass) and log(temperature) data to the data.frame
rmr_data$logM <- log(rmr_data$mass.kg)
rmr_data$logRMR <- log(rmr_data$RMR.mg.O2.h)

# calculate propensity scores considering method as the treatment and logM, T, therm as covariates:

# fit a propensity score model using nearest neighbor matching
ps_model <- matchit(method ~ logM, data = rmr_data, method = "nearest")
summary(ps_model)

# extract the matched dataset and fit the model again
matched_data <- match.data(ps_model)
mod_matched <- lm(logRMR ~ logM + T + therm + method, data = matched_data)
summary(mod_matched) # method does not have sig. effect.

# save rmr_data to file
save(rmr_data, file = "rmr_data.rda")

# save tree to file
save(tree, file = "tree.rda")
```

## Fit the RMR model

```{r}
# construct covariance matrix of species (Hadfield & Nakagawa, 2010).
A <- ape::vcv.phylo(compute.brlen(tree, method = "Grafen"))

# adding species as a factor into our dataset
rmr_data$spp_name <- factor(rmr_data$spp)

# Many species include serveral masurements. Thus, we use a bayesian models.
# General STAN specifications 
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
parallel::detectCores()

# MODEL
# We use a bayesian approach to study the relationship
# across species (phylogeny) while taking into considration
# intraspaecific variation and multiple observations for
# several species.


# set model priors
priors = c(
  prior(normal(0,10), "b"),
  prior(normal(0,50), "Intercept"),
  prior(student_t(3,0,20), "sd"),
  prior(student_t(3,0,20), "sigma")
)



```

Run the Bayesian model fit.


```{r}
# run Bayesian model including phylogenetic tree
set.seed(23); mod <- brm((logRMR)  ~ logM + T + therm # fixed coefficients
  + (1 | gr(spp, cov = A)) + (1| spp_name), # random factors
  family = gaussian(), 
  data2 = list(A = A),
  prior = priors,
  sample_prior = TRUE, 
  chains = 2, 
  cores = 16, # cores to be used 8 in mac pro
  iter = 3000, 
  warmup = 500,
  control = list(adapt_delta = 0.99, max_treedepth = 15), # Improve convergence
  save_pars = save_pars(all = TRUE),
  data = rmr_data)

# save the model outputs to file
save(mod, file = "mod.rda")
```

Check model convergence and summarise the output

```{r}

# get coefficients
summary(mod) 
coef_tab <- as.data.frame(summary(mod)$fixed)
round(coef_tab, 4)


# check model convergence 

par(oma = c(4,4,4,4))
plot(mod)
pp_check(mod)

# get bayesian p-value
p_map(mod)

pars <- brms::fixef(mod)[,1]

# save the estimated parameter values for the RMR regression
save(pars, file = "pars.rda")
```


## Define other key parameters

These parameters are shared by the figure generations files and are defined here and saved locally.

```{r}
# parameters for the estimation of k 
# these values derived from a regression that 
# excludes the two big planktivores but includes
# the big white shark.
kpars <- c(0.00174, -0.63180) 
save(kpars, file = "kpars.rda")

```


